### 2020-09-10 drafts/draft0.Rmd
added maths

---
title: "The increasing cost of happiness"
author: "Dr Richard Morris"
date: "Updated: `r Sys.Date()`"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(bayesplot)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")

#### To do ####
# https://www.psychologytoday.com/au/blog/how-happiness/201409/does-money-really-buy-happiness
# 
# more people could a achieve a level of wealth at which happiness no longer increased
```
***

## Abstract
A fundamental question for governments and people is how much happiness does a dollar buy? The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth. While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time. We tested the relationship between money and both happiness and life satisfaction using household economic data from Australia between 2002-2016. We discovered no change point with wealth existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living). Results such as these suggest we need to understand why the cost of happiness has increased, and to consider whether policy-goals to improve wellbeing by increasing wealth are feasible.

<br>

### Introduction
***
A fundamental question for the psychology and economics of wellbeing is just how much happiness does a dollar buy? Increasing wealth is commonly associated with increasing happiness, however a point at which wealth no longer increases happiness has also been [widely observed](https://www-sciencedirect-com.ezproxy2.library.usyd.edu.au/science/article/pii/S0167487007000694){target="_blank"} [@clark2008relative; @dolan2008we; @easterlin1974growth]. Given that a central goal of nations and governments is to improve wealth under the assumption that wealth always increases wellbeing, challenges to this notion have far reaching consequences [@frijters2020happy].

For instance, one [survey of 1,000 Americans](https://www.pnas.org/content/107/38/16489.full){target="_blank"}, conducted in 2010, concluded that money does make us happier – but only up to a certain point. The findings showed that self-reported levels of happiness increased with household wealth up to \$75,000 a year. But after that, increasing amounts of money had no further effect on happiness [@kahneman2010high]. Results such as these have reinforced the view among psychologists and economists that the relationship between wealth and measures of subjective wellbeing, such as happiness, increases linearly up to a point after which there is little gained by further increments in wealth [@clark2018four; @dolan2008we].

Subjective wellbeing is not a unitary entity [@diener2017findings]; studies typically distinguish between _life satisfaction_, the cognitive appraisal of one's own accomplishments, and _happiness_, one's prevailing affective state or emotional mood. Money can have different effects on each. For instance, we have recently reported that positive life events, such as a major financial windfall, can have a greater impact on an individual's satisfaction than their happiness [@kettlewell2020differential]. However these effects were observed as a change within individuals, and the change was transitory as both satisfaction and happiness returned to baseline after two years [see also @frijters2011life]. Across individuals, wealth also has a distinct relationship with happiness and satisfaction as well. While Kahneman et al observed the association between household wealth and happiness plateaud after $75,000, they also reported the association between wealth and life satisfaction was consistently positive. Indeed, several researchers have observed a difference between two questions that are often used in surveys of subjective well-being: “How satisfied are you with your life?” and “How happy are you these days?” [@veenhoven2006rising; @howell2008relation]. The common conclusion is that wealth is more strongly related to satisfaction than to happiness.

The relationship between wealth and happiness and satisfaction are relevant for the growing popularity of using subjective well-being as a policy guide for government and agencies. If there are distinct effects of wealth on happiness and satisfaction then the question of which is better suited to assess human welfare and to guide policy seems critical. The observation that happiness plateaus with increasing wealth while satisfaction keeps rising acquires some importance in this context. Not everyone will agree that enhancing the happiness experienced by those who are already quite happy is a legitimate policy objective.

And while research since 2010 has revealed variations in the plateau or _change point_ due to [world region, gender, and education](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} [@jebb2018happiness], there has been no investigation as to whether the relationship between wealth and happiness is changing over time [although see Fig. 4 in @clark2008relative; for a speculative mechanism]. The value of money of course changes over time with inflation, however the distribution of wealth has also changed over the last 30 years in many OECD countries including Australia (as the rich got richer faster than the poor, ref needed). Such changes in the distribution of wealth are also likely to change the relationship with wellbeing when it is governed by relative wealth rather than absolute wealth [@clark2008relative].

We used household economic panel data from Australia (HILDA) to provide the first investigation of whether changes in wealth and wellbeing have changed over the last 16 years (2002-2018). We distinguished between satisfaction and happiness as different components of subjective wellbeing, and evaluated how each varies with household wealth. Economic security is better represented by household wealth, since members of the same household share expenses as well as risks; i.e., they can provide a direct and immediate support network when financial shocks occur. The other major studies also used household after-tax income as the indicator of wealth and economic security, and so we follow the same standard here as well.

After adjusting for age, gender, and education level, we find that satisfaction increases linearly with household wealth, but the same purchasing power has diminishing returns on happiness. We also discover that the location of the inflection point has increased in real dollar value since 2002.

<br>

#### Wealth
We have a variety of income & wealth information in HILDA, most of it collected annually. However the household net worth is collected every four years. In the present report we use [hifdip](https://www.online.fbe.unimelb.edu.au/HILDAodd/KWCrossWaveCategoryDetails.aspx?varnt=hifdip){target="_blank"} Household financial year disposable regular income ($). The sum across all household members of financial year gross regular income minus taxes on financial year gross regular income. Wealth variables were provided by the University of Melbourne as imputed, and with weighted topcodes to preserve anonymity. _Nb_. We also examined household and personal gross wages and disposable income (not shown).

We removed individuals with values above the threshold indicated for top-coding. Household wealth was adjusted for the number of people in each household in the same manner as Headey, Muffels & Wooden, 2004; i.e., income divided by the square root of household size (Buhmann et al, 1988; Coulter et al, 1992).

<br>

#### Subjective Wellbeing
There are a variety of variables related to subjective well-being collected annually, but the two we used here matched the variables we used in our previous paper:

1. `losat` Life satisfaction - How satisfied are you with your life (0 to 10)
2. `gh9a` to `gh9i` Affective wellbeing - Sum of SF-36 item 9 (mental health and vitality)

_Note_. We also examined `ghmh` (the 5-item subset of mental health questions, sometimes called the **MHi-5**, from the SF-36), and the patterns were almost identical to our sum of 9a to 9i.

<br>

#### Modelling and model-selection

_Modelling_  
To initially describe the relationship between wealth and wellbeing, we modelled the data from each year in three different ways: 1) a simple linear model of the form y<sub>*i*</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>*x*<sub>*i*</sub> + *&epsilon;*<sub>*i*</sub>; 2) a cubic b-spline model of the form *B*(*x*) = *&Sigma;&beta;*<sub>*i*</sub>*B*<sub>*i,n*</sub>(*x*); and 3) a piecewise (broken-stick) model of the form y<sub>*i*</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;)(*x*<sub>*i*</sub> ≤ &omega;) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;)(*x*<sub>*i*</sub> ≤ &omega;).

These three models were chosen to represent varying degrees of complexity. The linear model was obviously the simplest relationship that can exist between wealth and wellbeing. The cubic B-spline was a complex model (more degrees of freedom) and chosen because it has been recently used to describe the maximum point at which wealth no longer increases wellbeing [@jebb2018happiness]. Following Jebb et al, we restricted the model to 4, 5 and 6 knots and show each instantiation here. The piecewise model was chosen as the simplest extension of a linear model to identify the change point at which wellbeing no longer increases linearly with wealth.

_Change point estimation_  
To determine whether the relationship between wellbeing and wealth contained an inflection point ("change point"), we modelled the relationship between wealth and wellbeing across individuals using the piecewise model described above, and estimated the posterior probability of the changepoint location using a Bayesian approach:

Let y<sub>*i*</sub> ~ *N*(*&mu;*<sub>*i*</sub>, *&sigma;*<sup>2</sup><sub>*y*</sub>)

*&mu;*<sub>i</sub> = *&beta;*<sub>0</sub> + *&beta;*<sub>1</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> ≤ &omega;<sub>*t*</sub>) + *&beta;*<sub>2</sub>(*x*<sub>*i*</sub> - &omega;<sub>*t*</sub>)(*x*<sub>*i*</sub> > &omega;<sub>*t*</sub>)  \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  (1)

Where *x*<sub>i</sub> and *y*<sub>i</sub> was an individuals' household wealth (\$) and wellbeing, respectively; and &omega;<sub>*t*</sub> was the changepoint in $ for each year (*t*).

The location of the changepoint was sampled randomly from a uniform prior over the resticted range of wealth values, and the posterior probability of the location for each year is presented.

<br>

_Model selection_  
To determine whether wellbeing was a linear or non-linear (e.g., piecewise) function of wealth, we compared the linear and piecewise model fits using the [Widely Applicable Information Criterion (WAIC)](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}. The WAIC is the log-posterior predictive density plus a penalty proportional to the variance in the posterior distribution. Thus it provides an approximation of the out-of-sample deviance that converges to the cross-validation approximation in a large sample, with a penalty for the effective number of parameters (degrees of freedom). For this reason is it useful to compare two models of varying complexity.

WAIC was defined as:

WAIC = -2(lppd - *p*~WAIC~)

Where lppd (log pointwise predictive density) is the total across observations of the log of the average likelihood of each observation, and *p*~WAIC~ is the effective number of free parameters determined by the sum of the variance in log-likelihood for each observation (*i*).

For visualization purposes only, due to the large number of individual data points in each year, we grouped individuals into 10 equal-sized subgroups according to their wealth decile ($). Thus in each plot, the mean wealth and mean wellbeing score for each decile is presented, rather than each individual data point. Note that the line-of-best-fit and 95% confidence intervals from the broken stick regression model _of all individuals_ is shown in overlay.

```{r funtimes}
source("../src/happyfun.r")
```

<br>

#### Covariates

Age (and age^2^), gender, and education were included as covariates, and students were removed.

```{r read_write_covs}
happywealth_covs <- read_rds("../results/happywealth_covs.rds")
```

<br>

### Results
***

#### Household wealth on satisfaction (cognitive wellbeing)
```{r losat_hwnwip, fig.height=11}
library(patchwork)

p1 <- happywealth_covs %>%
  plot_linear(losat, adj_hifdip)

p2 <- happywealth_covs %>%
  plot_inflection_cubic(losat, adj_hifdip)

p3 <- happywealth_covs %>%
  plot_inflection_piecewise(losat, adj_hifdip)

(p1 + p2 + p3) +
  plot_annotation(title = "Effect of wealth on satisfaction")
```

<br>

**Key points**

- Generally the relationship between satisfaction and household wealth appeared linear
- The linear relationship between satisfaction and wealth became weaker (less steep) over years
- There is no satiety point (horizontal inflection) in satisfaction before the highest decile in any year
- Any inflection point in satisfaction was as likely to be convex as concave

<br>

#### Household wealth on happiness (affective wellbeing)
```{r gh9_hwnwip, fig.height=11}
library(patchwork)

p1 <- happywealth_covs %>%
  plot_linear(gh9, adj_hifdip)

p2 <- happywealth_covs %>%
  plot_inflection_cubic(gh9, adj_hifdip)

p3 <- happywealth_covs %>%
  plot_inflection_custom(gh9, adj_hifdip, knots = c(32, 46, 52.5, 86, 124.5))

(p1 + p2 + p3) +
  plot_annotation(title = "Effect of wealth on happiness")
```

<br>

**Key points**

- The linear relationship between happiness and wealth became weaker (less steep) over years
- There was no happiness satiety point (horizontal inflection) before the final decile in any year
- A concave inflection point with happiness occurred in each year, and appeared to shift right with time.

<br>

##### Bayesian change point results

Shaded areas to the right of the vertical dotted line are credibly (95%) larger than the base year (2002).

```{r happy_change_time, fig.width = 4.9, fig.height = 9}
hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full.RDS")

as.data.frame(hifdip_gh9_fixed_fit) %>%
  select(b_alpha_Intercept, b_alpha_year2006, b_alpha_year2010, 
         b_alpha_year2014, b_alpha_year2018) %>%
  transmute(
    fy2002 = b_alpha_Intercept,
    fy2006 = mean(b_alpha_Intercept) + b_alpha_year2006,
    fy2010 = mean(b_alpha_Intercept) + b_alpha_year2010,
    fy2014 = mean(b_alpha_Intercept) + b_alpha_year2014,
    fy2018 = mean(b_alpha_Intercept) + b_alpha_year2018
  ) %>%
  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> posteriors

color_scheme_set("red")
mcmc_areas(
  posteriors,
  pars = rev(colnames(posteriors)),
  prob = 0.95, # 80% intervals
  prob_outer = 1, # 99%
  point_est = "mean"
  ) +
  theme_test() +
  geom_vline(aes(xintercept = mean(posteriors$fy2002)), color = "gray", linetype = 3) +
  labs(
    subtitle = "Posterior change points (mean & 95% distribution)",
    x = "Household wealth ($000s)")
```

<br>

**Key points**

- The change point for happiness (affective wellbeing) shifts to the right over time and occurs at greater household wealth levels.

<br>

#### Change points over time

The mean change point location has moved further right (higher, faster) than the median household wealth over time:

```{r, table1, ft.align="left"}
library(knitr)

happywealth_covs %>%
  select(year, xwaveid, adj_hifdip) %>%
  mutate(
    year = paste0("fy", year),
    adj_hifdip = adj_hifdip/1000) %>%
  left_join(
    gather(
      summarise_all(posteriors, list(median)), year, change_point
    )) %>%
  group_by(year) %>%
  summarise(
    `median ($000s)` = round(median(adj_hifdip), 2),
    `change point ($000s)` = round(mean(change_point), 2),
    `%` = round(mean(adj_hifdip > mean(change_point)), 2)
  ) %>%
  kable(format = "html", 
        caption = "Household wealth median, change point and percentile")
```

<br><br>

### Conclusions
***

The main conclusions of this report so far are:

- The relationship between subjective wellbeing and wealth is positive but happiness and satisfaction have different (positive) relationships. Satisfaction increases linearly with wealth while happiness increases rapidly up to a point, after which further increments in wealth produce less change - this is the **cost point** of happiness. So money buys happiness when economic security is scarce, but not so much once the cost point is reached.
- The cost point after which money provides decreasing returns in happiness is increasing over time. In 2002, the cost point of happiness represented a 13% increase over median wealth, while in 2018 it represented a 67% increase over median wealth.
- According to @frijters2020happy, coming up with a consensus estimate for income to translate into welbeing features high on the wellbeing research agenda. However evidence of a linear relationship between wellbeing and wealth calls into question the need to supplant traditional measures of economic success such as GDP with measures of wellbeing, such as "life satisfaction" surveys.

<br>

A study reported in 2013 challenged the idea that the positive effect of money plateaus [@stevenson2013subjective]. After comparing life satisfaction levels in both rich and poor countries, and rich and poor people within a country – with “rich” being defined as an income greater than $15,000 per person – Stevenson & Wolfers concluded: “The relationship between wellbeing and income … does not diminish as income rises. If there is a satiation point, we are yet to reach it.”

Since 2013, the distinct effect of wealth and happiness has also been challenged. [A survey of 1.7 million people representing 164 countries over 12 years (2005-2016) in the Gallup World Poll](https://www-nature-com.ezproxy2.library.usyd.edu.au/articles/s41562-017-0277-0){target="_blank"} reported money was associated with no further improvement in happiness after \$60-75,000, _nor any improvement in satisfaction_ after $95,000 [@jebb2018happiness]. Thus while the location differed, money had inflected effects with both satisfaction and happiness. On the other hand, in a nationally representative sample of 44,000 adult Americans over 44 years in the General Social Survey (GSS) happiness continued to increase with money [@twenge2020expanding], implying no inflection point existed between money and happiness.

Both these reports provide contradictory results regarding the nature of the relationship between wealth and wellbeing, and both are based on data spanning long periods of time (12-44 years). Perhaps the different results are due to differences in the constructs being measured, or perhaps it is due to different effects of time between studies.

### References