### 2021-02-08 src/supplementary.Rmd

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(flextable)
library(gtsummary)
library(bayesplot)
library(patchwork)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")
```
\  
\

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

\

**Corresponding author:**  
Professor Nick Glozier  
Faculty of Medicine and Health,   
University of Sydney,  
NSW 2050,  
Australia  
email: nick.glozier@sydney.edu.au

\

Draft: `r format(Sys.time(), '%d %B, %Y')`  
Supplementary tables:      4  
Supplementary figures:     3

**keywords:** Subjective wellbeing, household wealth, HILDA

\

***

## Supplementary Material

<br>

#### Sample characteristics

The broad demographic characteristics of the sample at each fourth year wave are presented below in Table S1.

```{r table_1, ft.align="left"}
happywealth <- read_rds("../data/happywealth.rds")

happywealth %>%
  filter(year %in% c(2002, 2006, 2010, 2014, 2018)) %>%
  filter(!student, !top_hifdip, (!is.na(losat)|!is.na(gh9))) %>%
  mutate(
    Sex = if_else(male, "Male", "Female"),
    Education = fct_collapse(
      edu,
      `High school or less` = c("None", "In school", "Highschool"),
      Graduate = c("Grad")
      ),
    Workforce = case_when(
      workforce & !unemployed ~ "Employed",
      unemployed ~ "Unemployed",
      TRUE ~ "Not in labour force"
      ),
    `Relationship status` = case_when(
      single ~ "Single",
      coupled ~ "Married",
      TRUE ~ "Divorced/Widow"
      ),
    Workforce = fct_relevel(Workforce, "Employed", "Unemployed"),
    `Relationship status` = fct_relevel(`Relationship status`, "Single", "Married")
    ) %>%
  select(
    year, 
    `Life satisfaction` = losat, `Happiness` = gh9,
    Sex, `Age (years)` = age, Education, Workforce, `Relationship status`, 
    `Chronic illness` = chronic, SEIFA, `Household size` = hhsize,
  ) %>% 
  tbl_summary(
    by = year,
    missing = "no",
    statistic = 
      list(all_continuous() ~ "{mean} (±{sd})",
           all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "(Missing)" 
    ) %>%
  modify_header(stat_by = "{level}\nN = {style_number(n)}") %>%
  as_flex_table() %>%
  set_caption(caption = "Table S1. Demographic characteristics of the sample") %>%
  align(part = "body") %>%
  valign(valign = "top", part = "header") %>%
  fontsize(size = 9.5, part = "all") %>%
  width(j = 1, width = 1.5) %>%
  width(j = 2:6, width = 1)
```

<br>

Average life satisfaction levels were very steady between 2002-2018, while average happiness score decreased slightly over the 17 years. The proportions of each sex and relationship status were stable over time, as were the average household size and SEIFA index. However age and chronic health conditions tended to increase over time, indicating a slight increase in the age of the panel over time (1.8 years) but not that of a cohort study which would be 16 years. Education levels (i.e., graduate percentage) also tended to increase over time, while changes in the workforce varied with economic circumstances.

<br><br>

#### Model Selection

We compared the posterior evidence of a linear- over a nonlinear- relationship between each wellbeing variable and income (i.e., WAIC~linear~ — WAIC~piecewise~). Thus a WAIC difference greater than zero indicated evidence for a linear relationship. A WAIC difference less than zero indicated evidence for a nonlinear (piecewise) relationship.

<br>

##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)

```{r complete_waic}
waic_gh9 <- read_rds("../results/model_comparisons_waic_R2_gh9.RDS")

waic_gh9 %>%
  select(model, year, elpd_diff, se_diff) %>%
  filter(elpd_diff != 0) %>%
  rowwise() %>%
  mutate(samples = list(rnorm(n = 1e5, 
                              mean = elpd_diff * 2, # for pw advantage
                              sd   = se_diff *  2)),
         year = paste0("AD", year)
         ) %>%
  select(year, samples) %>%
  group_by(year) %>%
  unnest(samples) %>% 
  mutate(sample_id = c(1:1e5)) %>%
  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
  select(-sample_id) -> d2

color_scheme_set("red")

p1 <- mcmc_intervals(x = d2,
           pars = colnames(d2),
           prob = 0.00,
           prob_outer = 0.95
           ) +
  scale_x_continuous(
    limits = c(-200, 150)
    ) +
  scale_y_discrete(labels = as.character(2002:2018)) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  theme_test() +
  labs(
    subtitle = "Happiness")

waic_losat <- read_rds("../results/model_comparisons_waic_R2_losat.RDS")

waic_losat %>%
  select(model, year, elpd_diff, se_diff) %>%
  filter(elpd_diff != 0, year > 2001) %>%
  rowwise() %>%
  mutate(samples = list(rnorm(n = 1e5, 
                              mean = elpd_diff * -2, # for linear advantage
                              sd   = se_diff *  2)),
         year = paste0("AD", year)
         ) %>%
  select(year, samples) %>%
  group_by(year) %>%
  unnest(samples) %>% 
  mutate(sample_id = c(1:1e5)) %>%
  pivot_wider(id_col = "sample_id", names_from = "year", values_from = "samples") %>%
  select(-sample_id) -> d1

color_scheme_set("blue")

p2 <- mcmc_intervals(x = d1,
  pars = rev(colnames(d1)),
  prob = 0.0,
  prob_outer = .95
  ) +
  scale_x_continuous(
    limits = c(-200, 150)
    ) +
  scale_y_discrete(labels = as.character(2002:2018)) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  theme_test() +
  labs(
    subtitle = "Satisfaction",
    x = "WAIC difference (95% CI)")
```

```{r figure_s1, fig.width=6.5, fig.height=7, dpi=300}
p1 / p2
```

```
Figure S1 legend: Differences in posterior evidence for a linear fit over a piecewise fit (WAIC) for happiness (red) and satisfaction (blue). The filled circle indicates the mean of the distribution and the horizontal bars represents the 95% credible interval. A difference greater than zero is support for the linear model and a difference below zero is support for the piecewise model.
```

<br>

The posterior evidence from model selection revealed the nonlinear (piecewise) fit of _happiness_, but not _satisfaction_, was credibly superior to a linear fit in each year. Figure S1 shows the 95% credible interval of the difference in WAIC values for the linear over nonlinear fits for happiness was below zero with no overlap for each year, indicating consistent support for the nonlinear model. By contrast, the WAIC differences for the satisfaction models were generally positive, indicating the superiority of the linear fit for this dependent variable. Overall the posterior evidence suggests that happiness and satisfaction have distinct relationships with household income; satisfaction tends to increase linearly with income, while a change point exists in the relationship between happiness and household income.

<br><br>

#### Model predictions (2002-2018)

The posterior estimates of the piecewise model indicated that the relationship between happiness and income changed over time between 2002 and 2018. One implication of such a change is the disparity in happiness between income groups has increased. This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related. To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which fell below the change points between 2002-2018 (\$40K/yr); and another level which always remained above the change points over the same period (\$75K/yr). Figure S2 below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).

<br>

##### Figure S2. Happiness (SD units) at $40K/yr and $75K/yr from 2002-2018 
```{r figure_s2, fig.width=7, fig.height=5, dpi=300, eval=T}
path_to_gh9_fits <- list.files(
  path = "../results/incl_topcodes/with_pweights/",
  pattern = "^fit_nl_hifdip_gh9_4cov_pwt_.*.RDS$",
  full.names = TRUE
)

predicted_fits <- list()
for (pathtofile in path_to_gh9_fits) {
  fit <- read_rds(pathtofile)

  ans <- model.frame(fit) %>%
    summarise_all(median)

  newdat = data.frame(
    dollars = c(4, ans[, "dollars"], 7.5),
    age = rep(ans[, "age"], 3),
    age_sq = rep(ans[, "age_sq"], 3),
    male = as.logical(rep(ans[, "male"], 3)), 
    grad = as.logical(rep(ans[, "grad"], 3)),
    pweights = rep(ans[, "pweights"], 3)
  )

  df <- fitted(fit, newdata = newdat) %>% 
    as_tibble() %>%
    mutate(
      Dollars = c("$40K", "\u03bc", "$75K"),
      year = as.numeric(
        substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4)
        )
      )

  predicted_fits <- bind_rows(predicted_fits, df)
}

predicted_fits %>%
  select(Estimate, Dollars, year) %>%
  filter(Dollars %in% c("$75K", "$40K")) %>%
  spread(Dollars, Estimate) %>%
  mutate(`$75K - $40K (∆)` = `$75K` - `$40K`) %>%
  gather(income, happiness, -year) %>%
  mutate(income = fct_relevel(income, "$40K", "$75K")) %>%
  ggplot(aes(x = year, y = happiness, color = income)) +
  geom_point(size = 4, alpha = 0.25) +
  geom_smooth(method = "loess", se = F, span = 3) +
  facet_wrap(~income) +
  labs(caption = "Source:     HILDA, University of Melbourne
                 Predicted happiness scores (SD unit) from piecewise regression on real household income", 
       x = "", y = "") -> p

source("ft_themes.r")

p +
  ft_theme() +
  ggthemes::scale_colour_wsj() +
  theme(legend.position = "none",
        plot.title = element_text(size = 19),
        strip.text = element_text(size = 12))
```

```{r alt_figure_s2, fig.width=7, fig.height=5, dpi = 300, eval=F}
happywealth %>%
  filter(!student) %>%
  filter(!top_hifdip) %>%
  select(year, xwaveid, gh9, income = hh_disp_inc_eq, weights) %>%
  filter(year %in% 2002:2018) %>%
  mutate(
    year = paste0("param_", year),
    income = income/1000) %>%
  mutate(
    year = extract_numeric(year),
    income_group = case_when(
      income > 35 & income < 45 ~ "$40K",
      income > 70 & income < 80 ~ "$75K"
    )) %>%
  group_by(year) %>%
  mutate(
    happiness = as.vector(scale(gh9)),
    ) %>%
  filter(!is.na(income_group)) %>%
  group_by(year, income_group) %>%
  summarise(
    Happiness = mean(happiness, na.rm = T)
    ) %>% 
  ungroup() %>%
  spread(income_group, Happiness) %>%
  mutate(`$75K - $40K (∆)` = `$75K` - `$40K`) %>%
  gather(key, val, `$40K`:`$75K - $40K (∆)`) %>%
  mutate(key = fct_relevel(key, c("$40K", "$75K"))) %>%
  ggplot(aes(x = year, y = val, color = key)) +
    geom_point(size = 4, alpha = 0.25) +
    geom_smooth(span = 3, se = F) +
    facet_wrap(~key) +
    labs(caption = "Source:     HILDA, University of Melbourne
                 Observed happiness scores (SD unit)", 
     x = "", y = "") -> p

p +
  ft_theme() +
  ggthemes::scale_colour_wsj() +
  theme(legend.position = "none",
        plot.title = element_text(size = 19),
        strip.text = element_text(size = 12))
```
```
Figure S2 legend: The difference (∆) in happiness between a household income of $40K per year and a household income of $75K per year has increased between 2002 and 2018.
```

<br>

Figure S2 shows the disparity in happiness between two fixed income levels, \$40K/year (below the change points) and \$75K/year (above the change points). In 2002, a household income of \$40K/year achieved a near average level of happiness relative to everyone else that year, however by 2018 happiness had declined to lower than average levels relative to the population. By contrast, a household income of \$75K/year enjoyed a higher than average level of happiness for the entire period. The difference (∆) in happiness between these two income levels doubled over the period. The increasing disparity in happiness between the rich and the poor implies that happiness has became more inequitable over the period.

<br><br>

***

##### Figure S3. Residual plots for piecewise happiness models
```{r figure_s3, fig.height=9, fig.width=9}
# https://www.flutterbys.com.au/stats/tut/tut7.2b.html
# resid = resid(data.brms)[, "Estimate"]
# fit = fitted(data.brms)[, "Estimate"]

resid_df <- readRDS("../results/residual_data.RDS")

resid_df %>%
  group_by(year) %>%
  slice_sample(n = 1000) %>% # to reduce overplotting
  ggplot(aes(x = fitteds, y = resids)) +
    geom_point(alpha = 0.1, size = 0.5) +
    facet_wrap(~year, scales = "free") +
    theme_test()
```

```
Figure S3 legend: The joint distribution between residual and fitted values shows no common patterns across years and appears clustered around zero in each case.
```

<br>

Examination of the residual plots from the piecewise models of happiness confirms the residuals were evenly distributed around zero and there was little indication of heterscedascity. Overall, there was little evidence that the piecewise models of happiness were not appropriate for the data.

<br><br>

#### Complete parameter estimates and fit statistics

There were _df_ = 59,876 total observations. 4000 samples (post-warmup) were drawn using sampling(NUTS). For each parameter, Bulk_ESS and Tail_ESS are effective sample size (ESS) measures, and Rhat is the potential scale reduction factor on split chains. In each case the Rhat apprached 1, indicating convergance. The smallest ESS was _n_ = 472, which is greater than a conservative estimate of the sufficient _n_ for estimation (*n* > 300).

The Bayes R^2^ = `r round(0.0403964581, 3)` ±`r round(0.001497172, 3)` (95%CI `r round(0.03672371,3)` — `r round(0.04262236, 3)`)

<br>

_Coefficient definitions_:

_alpha_ is the log value of wealth (x) at the changepoint  
_b0_ is the intercept denoting the expected value of wellbeing (y) at the changepoint  
_b1_ is the linear slope before the changepoint  
_b2_ is the linear slope after the changepoint  
_b3_ is the coefficient of education (graduate)  
_b4_ is the coefficient of age  
_b5_ is the coefficient of age^2^  
_b6_ is the coefficient of sex (male)

<br>

```{r gh9_pw_summary, ft.align="left"}
population_summary <- tibble()
for (pathtofile in path_to_gh9_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(
      Year = substring(pathtofile, nchar(pathtofile)-7, nchar(pathtofile)-4)
      ) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = str_remove(Parameter, "_Intercept")) %>%
  arrange(Parameter, desc(Year)) %>%
  filter(Year > 2001) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("b0", 
                           "b2", 
                           "b4", 
                           "b6"), 
     bg = "grey90", 
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S1. Piecewise parameter estimates for happiness") %>%
  autofit()
```

<br>

```{r gh9_lm_summary, ft.align="left"}
path_to_gh9_lm_fits <- list.files(
  path = "../results",
  pattern = '^fit_lm_hifdip_gh9_cov_.*.RDS$',
  full.names = TRUE
)

population_summary <- tibble()
for (pathtofile in path_to_gh9_lm_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 34, 37)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = fct_relevel(Parameter, "Intercept", "dollars", "grad")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("dollars",
                           "age",
                           "male"),
     bg = "grey90",
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S2. Linear parameter estimates for happiness") %>%
  autofit()
```

<br>

```{r losat_pw_summary, ft.align="left"}
path_to_losat_fits <- list.files(
  path = "../results",
  pattern = '^fit_cp_hifdip_losat_cov_.*.RDS$',
  full.names = TRUE
)

population_summary <- tibble()
for (pathtofile in path_to_losat_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 36, 39)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = str_remove(Parameter, "_Intercept")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("b0", 
                           "b2", 
                           "b4", 
                           "b6"), 
     bg = "grey90", 
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S3. Piecewise parameter estimates for satisfaction") %>%
  autofit()
```

<br>

```{r losat_lm_summary, ft.align="left"}
path_to_losat_lm_fits <- list.files(
  path = "../results",
  pattern = '^fit_lm_hifdip_losat_cov_.*.RDS$',
  full.names = TRUE
)

population_summary <- tibble()
for (pathtofile in path_to_gh9_lm_fits) {
  fit_summary <- read_rds(pathtofile) %>%
    summary()

  sigma_summary <- fit_summary$spec_pars %>%
    as_tibble(rownames = "Parameter")

  fixed_summary <- fit_summary$fixed %>%
    as_tibble(rownames = "Parameter") %>%
    bind_rows(sigma_summary) %>%
    mutate(Year = substring(pathtofile, 34, 37)) %>%
    select(Parameter, Year, everything())

  population_summary <- bind_rows(population_summary, fixed_summary)
}

population_summary %>%
  mutate(Parameter = fct_relevel(Parameter, "Intercept", "dollars", "grad")) %>%
  arrange(Parameter, desc(Year)) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Parameter %in% c("dollars",
                           "age",
                           "male"),
     bg = "grey90",
     part = "body") %>%
  bg(i = ~Parameter %in% c("sigma"), 
     bg = "#b3cde0", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S4. Linear parameter estimates for happiness") %>%
  autofit()
```

<br><br>

#### Additional non-linear fits

We tested the relationship between wealth and each wellbeing variable (happiness and satisfaction) using cubic **B-spline models**. A cubic B-spline was the most complex model (more degrees of freedom), and included here because it has been recently used to describe the maximum point at which wellbeing no longer increases with wealth [i.e., "satiety", @jebb2018happiness]. Following Jebb et al [-@jebb2018happiness], we restricted the cubic B-spline model to *n* = 4, 5 and 6 knots and display wealth on a log scale.

<br>

##### Figure S5. Cubic spline of household wealth on satisfaction and happiness
```{r cubic_plot, fig.height=9, fig.width=11}
source("../src/happyfun.r")
happywealth <- read_rds("../data/happywealth.rds")

p1 <- happywealth %>%
  filter(!student, !top_hifdip) %>%
  rename(Satisfaction = losat, `Household income` = hh_disp_inc_eq) %>% 
  plot_inflection_cubic(Satisfaction, `Household income`) +
  scale_x_continuous(
    # trans = "log", 
    # breaks = c(5, 10, 20, 40, 80, 160),
    limits = c(5, 170))

p2 <- happywealth %>%
  filter(!student, !top_hifdip) %>%
  rename(Happiness = gh9, `Household income` = hh_disp_inc_eq) %>% 
  plot_inflection_cubic(Happiness, `Household income`) +
  scale_x_continuous(
    # trans = "log", 
    # breaks = c(5, 10, 20, 40, 80, 160),
    limits = c(5, 170))

p1 / p2
```

<br>

The cubic B-spline models failed to reveal a complete satiety point (horizontal inflection) in satisfaction before the highest decile in any year. Also note that despite using a log scale, we do not observe a clear, regular logarithmic function with either wellbeing measure. Other researchers have argued that wealth represents a quantitative dimension subject to Weber's Law of perception [@kahneman2010high]. Under this view, similar proportional changes in income or wealth produce similar evaluations of satisfaction or happiness, rather than the absolute change. That is, a \$100 increase in wealth will produce a greater impact on someone earning a minimum wage than someone earning many orders of magnitude of that amount. However while it is true that wealth clearly represents a quantitative dimension, the effects of wealth observed here represent cross-sectional differences in the population rather than changes experienced by an individual over time. It is not clear the extent to which Weber's Law applies to cross-sectional or between-subject differences in quantitative dimensions such as wealth.

<br><br>

##### Figure S6. Prior predictive plot
(Not included)

```{r figure_s5, eval=F}
# prior_fit <- readRDS("../results/prior_nl_hifdip_losat_cov_2018.RDS")

# Check priors. The fixed estimates should be near zero, the fixed error should
# be ~3 respectively
# 
# summary(prior_fit)

# Extract prior predictions
# d <- data.frame(predict(prior_fit)) # takes a long time to run
# prior_predictions <- bind_cols(prior_fit$data, d)
# write_rds(prior_predictions, "../results/prior_prediction_data.RDS")

prior_predictions <- read_rds("../results/prior_prediction_data.RDS")

prior_predictions %>%
  mutate(
      dollar_bin = ntile(dollars, 10)
      ) %>%
    group_by(dollar_bin) %>%
    summarize(
      wellbeing = mean(Estimate, na.rm=T),
      wealth = mean(dollars, na.rm=T),
      n = n()
    ) -> df.deciles

prior_predictions %>%
  rename(wellbeing = Estimate,
         wealth = dollars) -> df.plot

# piecewise model
p0 <- ggplot(df.deciles, aes(x = wealth, y = wellbeing)) +
  geom_point() +
  geom_smooth(data = df.plot,
              aes(x = wealth, y = wellbeing,
                  fill = "#C79999", color = "#8F2727"),
              size = 0.5) +
  coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
                  ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))
                  ) +
  theme_test() +
  theme(legend.position = "none")

p1 <- ggplot(df.deciles, aes(x = wealth, y = wellbeing)) +
  geom_point() +
  geom_smooth(data = df.plot,
              aes(x = wealth, y = wellbeing, fill = "manual"),
              method = "lm",
              size = 0.5) +
  coord_cartesian(xlim = c(0, max(df.deciles$wealth)),
                  ylim = c(min(df.deciles$wellbeing), max(df.deciles$wellbeing))
                  ) +
  theme_test() +
  theme(legend.position = "none") +
  scale_fill_manual(values = "#b3cde0")

source("happyfun.R")

p2 <- prior_predictions %>%
  rownames_to_column(var = "xwaveid") %>%
  mutate(year = 2018,
         dollars = dollars*10000) %>%
  rename(Wellbeing = Estimate, `Household income` = dollars) %>% 
  plot_inflection_piecewise(Wellbeing, `Household income`, knots = 1)

p3 <- prior_predictions %>%
  rownames_to_column(var = "xwaveid") %>%
  mutate(year = 2018,
         dollars = dollars*10000) %>%
  rename(Wellbeing = Estimate, `Household income` = dollars) %>% 
  plot_inflection_cubic(Wellbeing, `Household income`)

p4 <- prior_predictions %>%
  rownames_to_column(var = "xwaveid") %>%
  mutate(year = 2018,
         dollars = dollars*10000) %>%
  rename(Wellbeing = Estimate, `Household income` = dollars) %>% 
  plot_linear(Wellbeing, `Household income`) +
  scale_fill_manual(values = "#b3cde0") +
  scale_color_manual(values = "#005b96")

p0 + p1

p2 + p3 + p4
```

<br><br><br>

##### Figure S7. Marginal effect of wealth on satisfaction by year (95% CI)  
(Not included)
```{r figure_s6_df, eval=F}
# Do not run

library(brms)

# import the fits
nl_losat_paths <- list.files(
  path = "../results/no_topcodes/",
  pattern = "fit_nl_hifdip_losat_cov_*",
  full.names = TRUE
)

list_fits = list()
for (f in nl_losat_paths) {
  fit <- readRDS(f)
  list_fits[str_extract(f, "[[:digit:]]+")] <- list(fit)
}

# turn each fit into a plot dataframe (this takes a few minutes)
plot_data <- map_dfr(.x = list_fits, 
                     .f = function(x) conditional_effects(x)[[1]],
                     .id = "Year")

saveRDS(plot_data, "../results/nl_marginal_effects_losat.RDS")
```
```{r, figure_s6, eval=F}
readRDS("../results/nl_marginal_effects_losat.RDS") %>%
  ggplot(aes(x = dollars, y = estimate__, fill = Year)) +
    geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.25) +
    geom_line(aes(color = Year)) +
    facet_wrap(~Year, scale = "free") +
    labs(y = "") +
    theme_test() +
    theme(legend.position = "none")
```

<br><br>

## References